---
title: "Practical Machine Learning - Week 3"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(magrittr)
library(data.table)
library(caret)
library(ggplot2)
library(rpart.plot)
library(MLmetrics)
```

## Tree Models

### Dataset

```{r}
data("iris")
table(iris$Species)
iris %>% setDT
```


```{r}
idx_train <- createDataPartition(iris$Species, p = 0.7, list = FALSE)
training <- iris[idx_train,]; dim(train)
testing <- iris[-idx_train,]; dim(test)
```

Plot the data by labels (species)

```{r}
ggplot(iris, aes(Petal.Length, Sepal.Width, color = Species)) +
  geom_point() +
  labs(title = "iris dataset")
```

### Model

```{r}
fit <- train(Species ~ ., method = "rpart", data = training)
rpart.plot(fit$finalModel, extra=104,
           branch.lty=3, shadow.col="gray", nn=TRUE)

# Evaluation on testing
y_pred <- predict(fit, type = "prob", newdata=testing)
table(testing$Species, y_pred)
# MLmetrics::MultiLogLoss(y_pred = y_pred, y_true = testing$Species)
```


## Bagging (Bootstrap aggregating)

**Method:**

1. Resample cases and recalculate predictions
2. Average or majority vote

**Notes:**

- Similar bias
- Reduce variance: averaging different predictors
- Most useful for non-linear functions

### Dataset

```{r}
library(ElemStatLearn)
library(data.table)
library(caret)
library(ggplot2)

data(ozone)
setDT(ozone)
ozone <- ozone[order(ozone),]
head(ozone)
```

### Bagged LOESS

by hand

```{r}
sample_idx <- caret::createResample(ozone$temperature, times = 10)

loess_ozone <- function(ozone, idx) {
  d <- ozone[idx,]; d[order(ozone)]
  loess0 <- loess(temperature ~ ozone, data = d, span = 0.2)
  out <- predict(loess0, data.table(ozone=1:155))
  data.table(ozone = 1:155, temperature = out)
}
res_list <- lapply(sample_idx, loess_ozone, ozone = ozone)
ll <- rbindlist(res_list, idcol = "group")

g <- ggplot(NULL, aes(ozone, temperature)) +
  geom_point(data = ozone, colour = "darkgrey")
g + 
  geom_line(aes(group = group), data = ll, colour = "grey") +
  geom_line(
  data = ll[, .(temperature=mean(temperature, na.rm = F)), ozone], 
  colour = "red"
)
```

by `caret` package

```{r}
tree_bag <- caret::bag(x = ozone[, .(ozone)], y = ozone$temperature, 
                       B = 10,
                       bagControl = bagControl(
                         fit = ctreeBag$fit,
                         predict = ctreeBag$pred,
                         aggregate = ctreeBag$aggregate
                       ))
new_data <- data.table(
  ozone[, .(ozone)],
  temperature = predict(tree_bag, newdata = ozone[, .(ozone)])
)
g + geom_point(data = new_data, colour = "blue")
```


## Random Forest

An extension to bagging

**Method:**

1. Resample cases and recalculate predictions (Bootstapp ing)
2. At each split, bootstrp variables (subset of variable to be used)
3. Average or majority vote

**Pros:**

1. Accuracy

**Cons:**

1. Speed low
2. Interpretability low
3. Overfitting

### Data
 
```{r}
library(ggplot2)
library(caret)
library(data.table)
data(iris); setDT(iris)

train_idx = createDataPartition(iris$Species, p = 0.7, list=FALSE)
training <- iris[train_idx,]
testing <- iris[-train_idx,]
```

### Random forest

```{r}
fit <- train(Species ~ ., 
             data = training, 
             method = "rf", 
             prox = TRUE) # for showing proximity
print(fit)

# Getting single tree
getTree(fit$finalModel, k=2)

# Class "centers"
centers <- classCenter(x = training[, .(Petal.Width, Petal.Length)],
            label = training[, Species],
            prox = fit$finalModel$proximity)
centers <- data.table(centers, Species = rownames(centers))
ggplot(NULL, aes(Petal.Width, Petal.Length, color = Species)) +
  geom_point(data = iris) +
  geom_point(data = centers, shape = 4, color = "black", size = 4)
```

### Cross validation

```{r}
set.seed(647)
myiris <- cbind(iris[1:4], matrix(runif(96 * nrow(iris)), nrow(iris), 96))
result <- rfcv(myiris, iris$Species, cv.fold=3)
with(result, plot(n.var, error.cv, log="x", type="o", lwd=2))
```

## Boosting

**Basic idea:**

1. Take a lot of (possibly) weak predictors
2. **Weight** them and add them
3. Get stronger predictor

## Model Based Prediction

Assumption to the distribution of the feature $x$ given the class $k$:

$$f_k(x) = Pr(X= x|Y = k)$$

- Linear Discriminant Analysis: assumes $f_k(x)$ is multivariate Gaussian with same covariances
- Quadratic Discriminant Analysis: assumes $f_k(x)$ is multivariate Gaussian with different covariances
- Naive Bayes: assumes independence between features


---

# Quiz

For this quiz we will be using several R packages. R package versions change over time, the right answers have been checked using the following versions of the packages.

- AppliedPredictiveModeling: v1.1.6
- caret: v6.0.47
- ElemStatLearn: v2012.04-0
- pgmm: v1.1
- rpart: v4.1.8

## Problem 1

```{r}
library(AppliedPredictiveModeling)
data(segmentationOriginal); setDT(segmentationOriginal)
library(caret)
library(data.table)
library(rpart.plot)
set.seed(125)

# 1. Subset the data to a training set and testing set based on the Case variable in the data set.
train_idx <- createDataPartition(segmentationOriginal$Case, p = 0.7, list=FALSE)

# 2. Set the seed to 125 and fit a CART model with the rpart method using all predictor variables and default caret settings.
training <- segmentationOriginal[Case == "Train"]
testing <- segmentationOriginal[Case == "Test"]

fit <- train(Class ~., data = training[, -c("Cell","Case")], method = "rpart")
predictors(fit)
print(fit$finalModel)
rpart.plot(fit$finalModel)
```

In the final model what would be the final model prediction for cases with the following variable values:

```{r}
newdata <- training[0]
newdata <- rbind(newdata, data.table(
  TotalIntench2 = c(23000, 50000, 57000, NA),
  FiberWidthCh1 = c(10, 10, 8, 8),
  PerimStatusCh1= c(2, NA, NA, 2),
  VarIntenCh4 = c(NA, 100, 100, 100)
), fill = TRUE)
predict(fit$finalModel, newdata = newdata, type = "class")
```


## Problem 3

```{r}
library(pgmm)
data(olive)
dim(olive)
head(olive)
olive <- olive[,-1]
treeModel <- train(Area ~ ., data=olive, method="rpart2")
treeModel
newdata <- as.data.frame(t(colMeans(olive)))
predict(treeModel, newdata) # 2.875
```

2.875. It is strange because Area should be a qualitative 
variable - but tree is reporting the average value of Area as 
a numeric variable in the leaf predicted for newdata

## Problem 4

```{r}
library(ElemStatLearn)
data(SAheart)
set.seed(8484)
train <- sample(1:dim(SAheart)[1],size=dim(SAheart)[1]/2,replace=F)
trainSA <- SAheart[train,]
testSA <- SAheart[-train,]
set.seed(13234)
logitModel <- train(chd ~ age + alcohol + obesity + tobacco + 
                      typea + ldl, data=trainSA, method="glm", 
                    family="binomial")
logitModel
missClass <- function(values,prediction){sum(((prediction > 0.5)*1) != values)/length(values)}
predictTrain <- predict(logitModel, trainSA)
predictTest <- predict(logitModel, testSA)
# Training Set Misclassification rate
missClass(trainSA$chd, predictTrain) # 0.2727273
# Test Set Misclassification rate
missClass(testSA$chd, predictTest) # 0.3116883
```

## Problem 5.

```{r}
library(ElemStatLearn)
library(randomForest)
data(vowel.train)
data(vowel.test)
head(vowel.train)
head(vowel.test)
dim(vowel.train) # 528  11
dim(vowel.test) # 462  11
vowel.train$y <- as.factor(vowel.train$y)
vowel.test$y <- as.factor(vowel.test$y)
set.seed(33833)
modelRf <- randomForest(y ~ ., data = vowel.train, importance = FALSE)
o <- order(varImp(modelRf), decreasing=T)
rownames(varImp(modelRf))[o]
# The order of the variables is:
#  x.2, x.1, x.5, x.6, x.8, x.4, x.9, x.3, x.7,x.10
```
