---
title: "Practical Machine Learning - Week 4"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  comment = "#>"
)
```

## Regularized Regression

1. Fit a regression model
2. Penalize (or shrink) large coefficients

**Pros:**

1. Can help with the bias variance tradeoff
2. Can help with the model selection

**Cons:**

1. Computational demanding on large datasets
2. Does not perform as well as Random Forests and boosting

In caret methods are:

- ridge
- lasso
- relaxo

## Combining Predictors

- Combine classifiers by averaging/voting
- Improve accuracy
- Reduce interpretability
- Boosting, bagging, random forests are variants on this theme

Approaches:

1. Bagging, boosting, random forests
    - Usually combine similar classifiers
2. Combining different classifiers
    - Model stacking
    - Model ensambling


### Model Stacking

Data

```{r}
library(ISLR)
library(data.table)
library(ggplot2)
library(caret)
library(magrittr)

data("Wage"); setDT(Wage)
Wage[, logwage := NULL]

# Create training set and validation set
idx_build <- createDataPartition(Wage$wage, p = 0.7, list = FALSE)
validation <- Wage[-idx_build]
train_idx <- createDataPartition(Wage[idx_build, wage], p = 0.7, list = FALSE)
training <- Wage[idx_build][train_idx]
testing <- Wage[idx_build][-train_idx]
```

Build two different models

```{r}
mod1 <- train(wage ~., method = "glm", data = training)
mod2 <- train(wage ~., method = "rf", data = training,
              trControl = trainControl(method = "cv"),
              number = 3)

pred1 <- predict(mod1, newdata = testing)
pred2 <- predict(mod2, newdata = testing)
```

Fit a model that **combine the predictors**

```{r}
pred_dt <- data.table(
  pred1,
  pred2,
  wage = testing$wage
)

comb_mod <- train(wage ~ ., method = "gam", data = pred_dt)
comb_pred <- predict(comb_mod, pred_dt)
```

Tesing errors

```{r}
# model 1
(pred1 - testing$wage)^2 %>% sum %>% sqrt()

# model 1
(pred2 - testing$wage)^2 %>% sum %>% sqrt()

# combined model
(comb_pred - testing$wage)^2 %>% sum %>% sqrt()
```

Predict on validation dataset

```{r}
pred_val_1 <- predict(mod1, validation)
pred_val_2 <- predict(mod2, validation)
pred_val_dt <- data.table(pred1 = pred_val_1, pred2 = pred_val_2)
comb_pred_val <- predict(comb_mod, pred_val_dt)

# Evaluation on validation
(comb_pred_val - validation$wage)^2 %>% sum %>% sqrt
```

## Forecasting (for time temporal data, spatial data)

- Goal: predict one or more observations in the future

```{r}
library(forecast)
library(quantmod)
GOOG <- getSymbols("GOOG", auto.assign = FALSE)
class(GOOG)
head(GOOG)
```

Summarize monthly and store as time series

```{r}
mGOOG <- to.monthly(GOOG)
goog_open <- Op(mGOOG)
ts1 <- ts(goog_open, frequency = 12)
plot(ts1, xlab = "Years+1") 
```

Time series decomposition

- Trend
- Seasonal
- Cyclic

```{r}
plot(decompose(ts1), xlab = "Years+1")
```

Training and test sets

```{r}
ts1_train <- window(ts1, start = 1, end = 5)
ts1_test <- window(ts1, start = 5, end = 7-0.01)
```

**Simple moving average**

$$Y_t = \frac{1}{2*k+1}\sum_{ j=-k }^{ k }{ y_{t+j} } $$

```{r}
plot(ts1_train)
lines(ma(ts1_train, 3), col = "red")
```

**Exponential Smoothing**

```{r}
ets1 <- ets(ts1_train, model = "MMM")
fcast <- forecast(ets1)
plot(fcast); lines(ts1_test, col = "red")
```

Accuracy

```{r}
accuracy(fcast, ts1_test)
```


## Unsupervised Prediction

K-means

```{r}
library(data.table)
library(caret)
library(ggplot2)
library(gridExtra)

data("iris")
setDT(iris)

# create training and test sets
inTrain <- createDataPartition(y=iris$Species,p=0.7, list=FALSE)
training <- iris[inTrain,]; testing <- iris[-inTrain,]

# perform k-means clustering for the data without the Species information
# Species = what the true clusters are
kMeans1 <- kmeans(training[, -"Species"], centers=3)
# add clusters as new variable to training set
training$clusters <- as.factor(kMeans1$cluster)

# plot clusters vs Species classification
p1 <- qplot(Petal.Width,Petal.Length,colour=clusters,data=training) +
    ggtitle("Clusters Classification")
p2 <- qplot(Petal.Width,Petal.Length,colour=Species,data=training) +
    ggtitle("Species Classification (Truth)")
grid.arrange(p1, p2, ncol = 2)
```


---

# Quiz

```{r}
library(ElemStatLearn)
library(data.table)
data(vowel.train); setDT(vowel.train)
data(vowel.test); setDT(vowel.test)

vowel.train[, y := as.factor(as.character(y))]
vowel.test[, y := as.factor(as.character(y))]
```

## 1.

Set the variable y to be a factor variable in both the training and test set. Then set the seed to 33833. Fit (1) a random forest predictor relating the factor variable y to the remaining variables and (2) a boosted predictor using the "gbm" method. Fit these both with the train() command in the caret package.

What are the accuracies for the two approaches on the test data set? What is the accuracy among the test set samples where the two methods agree?

```{r}
library(caret)
library(doMC)
registerDoMC(cores = 3)

set.seed(33833)

fit1 <- train(y ~., method = "rf", 
              trControl = trainControl(method = "cv"),
              data = vowel.train)
fit2 <- train(y ~., method = "gbm", data = vowel.train)

pred_rf <- predict(fit1, vowel.test)
pred_gbm <- predict(fit2, vowel.test)

# Accuracy
confusionMatrix(pred_rf, vowel.test$y) # RF Accuracy: 0.6060606
confusionMatrix(pred_gbm, vowel.test$y) # GBM Accuracy: 0.530303

# Vote by two models
pred <- data.table(pred_rf, pred_gbm, 
                   y = vowel.test$y, 
                   agree = pred_rf == pred_gbm)
# Agreement Accuracy
hit <- pred[agree == TRUE & pred_rf == y] %>% nrow
hit / pred[, sum(agree)] # Agreement Accuracy: 0.6569579
```

## 2.

Load the Alzheimer's data using the following commands

```{r}
library(data.table)
library(caret)
library(gbm)
library(AppliedPredictiveModeling)
library(doMC)
registerDoMC(cores = 3)

set.seed(3433)

data(AlzheimerDisease)
adData = data.frame(diagnosis,predictors)
inTrain = createDataPartition(adData$diagnosis, p = 3/4)[[1]]

training = adData[ inTrain,]
testing = adData[-inTrain,]
```

Set the seed to 62433 and predict diagnosis with all the other variables using a random forest ("rf"), boosted trees ("gbm") and linear discriminant analysis ("lda") model. Stack the predictions together using random forests ("rf"). What is the resulting accuracy on the test set? Is it better or worse than each of the individual predictions?

```{r}
set.seed(62433)

mod_rf <- train(diagnosis ~., method = "rf", data = training,
                trControl = trainControl(method = "cv"),
                number = 3)
mod_gbm <- train(diagnosis ~., method = "gbm", data = training)
mod_lda <- train(diagnosis ~., method = "lda", data = training)

stack_dt <- data.table(
  predict(mod_rf, testing),
  predict(mod_gbm, testing),
  predict(mod_lda, testing),
  diagnosis = testing$diagnosis
)

mod_stack <- train(diagnosis ~., method = "rf", data = stack_dt)

# Accuracy
confusionMatrix(predict(mod_rf, testing), testing$diagnosis)$overall # 0.792
confusionMatrix(predict(mod_gbm, testing), testing$diagnosis)$overall # 0.780
confusionMatrix(predict(mod_lda, testing), testing$diagnosis)$overall # 0.768
confusionMatrix(predict(mod_stack, stack_dt), testing$diagnosis)$overall # 0.80
```

## 3. 

Load the concrete data with the commands:

```{r}
library(data.table)
library(caret)
library(gbm)
library(AppliedPredictiveModeling)
library(doMC)
set.seed(3523)
library(AppliedPredictiveModeling)

data(concrete); setDT(concrete)
inTrain = createDataPartition(concrete$CompressiveStrength, p = 3/4)[[1]]
training = concrete[ inTrain,]
testing = concrete[-inTrain,]
```

Set the seed to 233 and fit a lasso model to predict Compressive Strength. Which variable is the last coefficient to be set to zero as the penalty increases? (Hint: it may be useful to look up ?plot.enet).

```{r}
set.seed(233)
fit <- train(CompressiveStrength ~ ., data=training, method="lasso")
fit
plot.enet(fit$finalModel, xvar="penalty", use.color=T) # Cement
```

## 4.

Load the data on the number of visitors to the instructors blog from here:

https://d396qusza40orc.cloudfront.net/predmachlearn/gaData.csv

Using the commands:

```{r}
library(lubridate) # For year() function below
library(data.table)
library(readr)
dat <- read_csv("https://d396qusza40orc.cloudfront.net/predmachlearn/gaData.csv")
setDT(dat)
dat[, X1 := NULL]
training = dat[year(dat$date) < 2012,]
testing = dat[(year(dat$date)) > 2011,]
tstrain = ts(training$visitsTumblr)
```

Fit a model using the bats() function in the forecast package to the training time series. Then forecast this model for the remaining time points. For how many of the testing points is the true value within the 95% prediction interval bounds?

```{r}
library(forecast)
fit <- bats(tstrain)

pred <- forecast(fit, level=95, h = dim(testing)[1])
pred_comb <- cbind(testing, as.data.table(pred))
setnames(pred_comb, 
         names(pred_comb), 
         gsub("[^\\w]", "_", names(pred_comb), perl = TRUE))
pred_comb[, in95 := (Lo_95 < visitsTumblr) & (visitsTumblr < Hi_95)]

# How many of the testing points is the true value within the 
# 95% prediction interval bounds?
prop.table(table(pred_comb$in95)) # 0.9617021
```

## 5.

Load the concrete data with the commands:

```{r}
library(data.table)
library(caret)
library(gbm)
library(AppliedPredictiveModeling)
library(doMC)
set.seed(3523)
library(AppliedPredictiveModeling)

data(concrete); setDT(concrete)
inTrain = createDataPartition(concrete$CompressiveStrength, p = 3/4)[[1]]
training = concrete[ inTrain,]
testing = concrete[-inTrain,]
```

Set the seed to 325 and fit a support vector machine using the e1071 package to predict Compressive Strength using the default settings. Predict on the testing set. What is the RMSE?

```{r}
library(e1071)
set.seed(325)
fit <- svm(CompressiveStrength ~., data=training)

pred <- predict(fit, testing)
acc <- accuracy(pred, testing$CompressiveStrength) # RMSE 6.715009
```


